---
title: "Case approach"
author: "Bárbara Barbosa"
date: "November 1, 2016"
output: pdf_document
---

Load packages and read file. Check out some data.
```{r}
library(randomForest)
library(MASS)
base <- read.csv('~/workspace/game_return/training_set.csv', na.strings = 'NULL')
summary(base)
```

Normalize variables, since `read.csv` reads columns as categorical variables, it's necessary to change them to numbers.

```{r}
normalize_vars <- function(base){
  base['revenue'] <- as.double(as.character(base['revenue'][[1]]))
  base['unit_count'] <-  as.double(as.character(base['unit_count'][[1]]))
  base['try_1'] <- as.integer(as.character(base['try_1'][[1]]))
  base['win_1'] <- as.integer(as.character(base['win_1'][[1]]))
  base['try_2'] <- as.integer(as.character(base['try_2'][[1]]))
  base['win_2'] <- as.integer(as.character(base['win_2'][[1]]))
  base['try_3'] <- as.integer(as.character(base['try_3'][[1]]))
  base['win_3'] <- as.integer(as.character(base['win_3'][[1]]))
  base['ttp'] <- as.integer(as.character(base['ttp'][[1]]))
  base['session_count'] <- as.integer(as.character(base['session_count'][[1]]))
  base
}

base <- normalize_vars(base)
```

Since there are some NAs and the different NAs do not belong to the same records, I changed some NAs to 0 and removed the `try` and `win` vars that are NAs.
```{r}
sapply(base, function(x) sum(is.na(x)))

remove_nas <- function(base){
  base <- base[!is.na(base['try_1'][[1]]),]
  base <- base[!is.na(base['try_2'][[1]]),]
  base <- base[!is.na(base['try_3'][[1]]),]
  base <- base[!is.na(base['win_1'][[1]]),]
  base <- base[!is.na(base['win_2'][[1]]),]
  base <- base[!is.na(base['win_3'][[1]]),]

  base['session_count'][is.na(base['session_count'])] <- 0
  base['ttp'][is.na(base['ttp'])] <- 0
  base['unit_count'][is.na(base['unit_count'])] <- 0
  base['revenue'][is.na(base['revenue'])] <- 0  
  base
}

base <- remove_nas(base)
```


Some simple exploratory analysis to know the data:
```{r}
barplot(table(base$country), main = 'Country')

barplot(table(base$device), main = 'Device')

plot(table(base$ttp), main = 'Total time played')

plot(base$ttp, base$session_count, main = 'Total time played in sessions')

barplot(table(as.double(base$revenue)), main = '#Revenue')

plot(base$try_1, base$win_1, main='Activity 1')

plot(base$try_2, base$win_2, main='Activity 2')

plot(base$try_3, base$win_3, main='Activity 3')
```

It looks like activity 3 is the least popular and activity 2 is the most popular. Also, the plots allowed me to verify my cleaning strategies.

Verify the try and game correlations. The plots already showed that they have a high correlation, it's just to verify.
```{r}
cor(base$try_1, base$win_1)
cor(base$try_2, base$win_2)
cor(base$try_3, base$win_3)
```

Variable creation. Since the both are high correlated,  I created the `win_rate`, that has more info from both variables. Since some values are 0, it is necessary to replace NaN to 0.
```{r}
create_win_rate <- function(base){
  base$win_rate_1 <- round((base['win_1'][[1]]/base['try_1'][[1]])*100)
  base$win_rate_1[is.nan(base$win_rate_1)] <- 0
  base$win_rate_1[is.infinite(base$win_rate_1)] <- 0

  base$win_rate_2 <- round((base['win_2'][[1]]/base['try_2'][[1]])*100)
  base$win_rate_2[is.nan(base$win_rate_2)] <- 0
  base$win_rate_2[is.infinite(base$win_rate_2)] <- 0

  base$win_rate_3 <- round((base['win_3'][[1]]/base['try_3'][[1]])*100)
  base$win_rate_3[is.nan(base$win_rate_3)] <- 0
  base$win_rate_3[is.infinite(base$win_rate_3)] <- 0
  base
}

base <- create_win_rate(base)
```


Verify if there is a cut on any win_range that separates the days better
```{r}
r_labels = c('0-10', '10-20', '20-30', '30-40', '40-50','50-60', '60-70', '70-80', '80-90', '90-100')
r_breaks = c(-Inf, 10, 20, 30, 40, 50, 60, 70, 80, 90, +Inf)

# activity 1

win_range <- cut(base$win_rate_1, breaks = r_breaks, labels = r_labels)

u = data.frame(as.factor(win_range), base$days_active)
mean_range <- c(mean(base$win_rate_1[base$days_active == 1]), mean(base$win_rate_1[base$days_active == 2]), mean(base$win_rate_1[base$days_active == 3]), mean(base$win_rate_1[base$days_active == 4]), mean(base$win_rate_1[base$days_active == 5]), mean(base$win_rate_1[base$days_active == 6]), mean(base$win_rate_1[base$days_active == 7]), mean(base$win_rate_1[base$days_active == 8]))

barplot(mean_range, main = 'Mean active day for each win range in activity 1', col=heat.colors(1), names.arg=c("1","2","3","4","5", "6", "7","8"))

#activity 2

win_range <- cut(base$win_rate_2, breaks = r_breaks, labels = r_labels)

u = data.frame(as.factor(win_range), base$days_active)
mean_range <- c(mean(base$win_rate_2[base$days_active == 1]), mean(base$win_rate_2[base$days_active == 2]), mean(base$win_rate_2[base$days_active == 3]), mean(base$win_rate_2[base$days_active == 4]), mean(base$win_rate_2[base$days_active == 5]), mean(base$win_rate_2[base$days_active == 6]), mean(base$win_rate_2[base$days_active == 7]), mean(base$win_rate_2[base$days_active == 8]))

barplot(mean_range, main = 'Mean active day for each win range in activity 2', col=heat.colors(1), names.arg=c("1","2","3","4","5", "6", "7","8"))

#activity 3

win_range <- cut(base$win_rate_3, breaks = r_breaks, labels = r_labels)

u = data.frame(as.factor(win_range), base$days_active)
mean_range <- c(mean(base$win_rate_3[base$days_active == 1]), mean(base$win_rate_3[base$days_active == 2]), mean(base$win_rate_3[base$days_active == 3]), mean(base$win_rate_3[base$days_active == 4]), mean(base$win_rate_3[base$days_active == 5]), mean(base$win_rate_3[base$days_active == 6]), mean(base$win_rate_3[base$days_active == 7]), mean(base$win_rate_3[base$days_active == 8]))

mean_range[is.infinite(mean_range)] <- 0

barplot(mean_range, main = 'Mean active day for each win range in activity 3', col=heat.colors(1), names.arg=c("1","2","3","4","5", "6", "7","8"))
```

Based on the graphs, I chose a cut using the day 1 `win_rate` (the smallest) to create the `high_win_rate` variable.

```{r}
create_high_win_rate <- function(base){
  base$high_win_rate_1 <- base['win_rate_1'][[1]] >= 60
  base$high_win_rate_2 <- base['win_rate_2'][[1]] >= 50
  base$high_win_rate_3 <- base['win_rate_3'][[1]] >= 20
  base
}

base <- create_high_win_rate(base)
```

Verify `session_count`, and create the variables `less_than_three_sessions` and `play_more_thirty_min`. Also verify if the player has more than 50% of wins in at least one activity.

```{r}
session_range <- cut(base$session_count, breaks = r_breaks, labels = r_labels)

u = data.frame(as.factor(session_range), base$days_active)
mean_range <- c(mean(base$session_count[base$days_active == 1]), mean(base$session_count[base$days_active == 2]), mean(base$session_count[base$days_active == 3]), mean(base$session_count[base$days_active == 4]), mean(base$session_count[base$days_active == 5]), mean(base$session_count[base$days_active == 6]), mean(base$session_count[base$days_active == 7]), mean(base$session_count[base$days_active == 8]))

mean_range[is.infinite(mean_range)] <- 0

barplot(mean_range, main = 'Mean active day for each session count', col=heat.colors(1), names.arg=c("1","2","3","4","5", "6", "7","8"))

base$less_than_three_sessions <- base$session_count < 3

#ttp

ttp_range <- cut(base$ttp, breaks = r_breaks, labels = r_labels)

u = data.frame(as.factor(ttp_range), base$days_active)
mean_range <- c(mean(base$ttp[base$days_active == 1]), mean(base$ttp[base$days_active == 2]), mean(base$ttp[base$days_active == 3]), mean(base$ttp[base$days_active == 4]), mean(base$ttp[base$days_active == 5]), mean(base$ttp[base$days_active == 6]), mean(base$ttp[base$days_active == 7]), mean(base$ttp[base$days_active == 8]))

barplot(mean_range, main = 'Mean active day for each tpp', col=heat.colors(1), names.arg=c("1","2","3","4","5", "6", "7","8"))

base$play_more_thirty_min <- base$ttp > 1800

base$more_than_half_wins <- ifelse( base$high_win_rate_1 > 0, TRUE, ifelse(base$high_win_rate_2 > 0, TRUE, ifelse(base$high_win_rate_3, TRUE, FALSE)))
```

Normalize `days_active` to be able to use as classification predictor

```{r}
base$more_than_one_day <- ifelse(base$days_active > 1, 1, 0)
```

Generate a train and validation samples. Train has 70% of observations and validation has 30%. Define a MSE function to validate the algorithms.

```{r}
set.seed(78)
sample_size <- floor(0.7 * length(base[,1]))
training_data_rows <- sample(seq_len(nrow(base)), size = sample_size)
train_data <- base[training_data_rows,]
test_data <- base[-training_data_rows,]
base_list <- list(train_data, test_data)
train_set <- base_list[[1]]
test_set <- base_list[[2]]

MSE = function(m, o){
  mean((m - o)^2)
}
```

Test stepwise selection forward and backwards to verify the best variables and generate a model (`best_fit`) with it. (MSE for test_set is 4.464)

```{r}
all_fit <- lm(days_active ~ more_than_half_wins + less_than_three_sessions + country + device + session_count * ttp + unit_count + revenue + try_1 + try_2 + try_3 + win_1 + win_2 + win_3 + win_rate_1 + win_rate_2 + win_rate_3 + high_win_rate_1 + high_win_rate_2 + high_win_rate_3, data = train_set)

step <- stepAIC(all_fit, direction="both")
step$anova # display results 

best_fit <- lm(days_active ~ less_than_three_sessions + country + device + session_count + ttp + unit_count + win_1 + win_2 + win_3 + win_rate_2 + high_win_rate_3 + session_count*ttp, data = train_set)
summary(best_fit)

test_set$best_days_predict <- round(predict(best_fit, test_set))

MSE(test_set$best_days_predict, test_set$days_active)
```

Final model - Linear regression model - Has the best MSE for the test set (4.443)

```{r}
fit <- lm(days_active ~ play_more_thirty_min + country + device + session_count * ttp + less_than_three_sessions + unit_count + high_win_rate_3 + win_rate_1 + win_rate_2 + win_rate_3, data = train_set)
summary(fit)

test_set$days_predict <- round(predict(fit, test_set))

test_set$days_predict[test_set$days_predict > 8] <- 8

MSE(test_set$days_predict, test_set$days_active)
```

Test randomForest for regression (regression forest) to verify the MSE. (the forest does not generate values for day 1 and greater than 5) The best model is the 2nd linear regression.

```{r}
###############
tree_fit <- randomForest(days_active ~ country + device + session_count + ttp + unit_count + revenue + try_1 + try_2 + try_3 + win_1 + win_2 + win_3, data = train_set, importance=TRUE, na.action = na.omit)
print(tree_fit)
round(importance(tree_fit), 2)
summary(tree_fit)

test_set$tree_days_predict <- round(predict(tree_fit, test_set))

MSE(test_set$tree_days_predict, test_set$days_active)
```

Classification random forest

```{r}
classification_fit <- randomForest(as.factor(more_than_one_day) ~ country + device + session_count + ttp + unit_count + revenue + try_1 + try_2 + try_3 + win_1 + win_2 + win_3 + win_rate_1 + win_rate_2 + win_rate_3, importance=TRUE, proximity=TRUE, data = train_set)

print(classification_fit)

summary(classification_fit)

test_set$greater_than <- predict(classification_fit, test_set)

true_table <- table(test_set$greater_than, test_set$more_than_one_day)

#acuracy
(true_table[1] + true_table[4])/(true_table[1] + true_table[2] + true_table[3] + true_table[4])
```

Generate responses for test file. First clean and add vars to data, then calculate the model response.

```{r}
test_base <- read.csv('~/workspace/game_return/test_set.csv', na.strings = 'NULL')

# normalize csv vars
test_base <- normalize_vars(test_base)
test_base <- remove_nas(test_base)

#create new vars
test_base <- create_win_rate(test_base)
test_base <- create_high_win_rate(test_base)
test_base$play_more_thirty_min <- test_base$ttp > 1800
test_base$less_than_three_sessions <- test_base$session_count < 3

test_base$days_active <- round(predict(fit, test_base))

test_base$days_active[test_base$days_active > 8] <- 8

test_base$greater_than <- predict(classification_fit, test_base)

final_file <- data.frame(test_base$user_id, test_base$days_active, test_base$greater_than)

write.csv(final_file, 'answer.csv')
```

